Pigeons.jl: Distributed sampling from intractable distributions

### 2024-05-21 paper/PT.tex
Add new cite and fix typo
New tweaks based on Alex's comments

\medskip
We emphasize again that Pigeons supports both shared memory 
parallelism (SMP) and distributed memory parallelism (DMP). However, in certain cases, it 
might be useful to run parallel tempering with very many chains or 
to run several instances of parallel tempering at once. In such cases, 
the DMP setup is the most natural. Our implementation in Pigeons ensures that the 
communication cost in this setup is minimal. In most cases, the 
communication cost is negligible compared to other computations performed on each process.

\medskip
The weak scaling results for both the DMP and SMP settings are presented in \cref{fig:PT_scaling}. 
We see that we obtain an approximately linear speedup as we increase the number of processes.

\begin{figure*}[t]
    \centering
    \begin{minipage}{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/speedup_distributed.pdf}
    \end{minipage}
    \begin{minipage}{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/speedup_parallel.pdf}
    \end{minipage}
    \caption{
        Weak scaling results for PT in both the distributed (MPI) and parallel settings 
        with approximate 95\% confidence intervals for the speedup.  
        The blue dotted line represents a hypothetical speedup equal to the 
        number of processes/threads requested. 
        The red dashed line indicates a speedup of one for reference.
        \textbf{Left:} Distributed setup. 
        \textbf{Right:} Parallel (shared) setup. 
    }
    \label{fig:PT_scaling}
\end{figure*}

>>>>

\medskip
The weak scaling results for both the distributed memory parallelism (DMP) and 
shared memory parallelism (SMP) settings are presented in \cref{fig:PT_scaling}. 
We see that we obtain an approximately linear speedup as we increase the number of processes.

\begin{figure*}[t]
    \centering
    \begin{minipage}{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/speedup_distributed.pdf}
    \end{minipage}
    \begin{minipage}{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/speedup_parallel.pdf}
    \end{minipage}
    \caption{
        Weak scaling results for PT in both the distributed (MPI) and parallel settings 
        with approximate 95\% confidence intervals for the speedup.  
        The blue dotted line represents a hypothetical speedup equal to the 
        number of processes/threads. 
        The red dashed line indicates a speedup of one for reference.
        \textbf{Left:} Distributed setup. 
        \textbf{Right:} Parallel (shared memory) setup. 
    }
    \label{fig:PT_scaling}
\end{figure*}

\subsection{Relevance of DMP versus SMP}

We emphasize again that Pigeons supports both shared memory 
parallelism (SMP) and distributed memory parallelism (DMP). However, in certain cases, it 
might be useful to run parallel tempering with very many chains or 
to run several instances of parallel tempering at once. In such cases, 
the DMP setup is the most natural if the requested number of chains is very large. 
Another notable example of the benefit of DMP is for the multi-GPU use case. 
For instance, in some situations the likelihood evaluation of a single chain 
requires its own GPU (e.g., for high-resolution black hole imaging, which relies 
on running a Fast Fourier Transform on a GPU). 
Additionally, DMP can be useful when the user's code is not thread safe. 
In such cases, concurrency can still be achieved via DMP.
Our implementation in Pigeons ensures that the 
communication cost in the DMP setup is minimal. In most cases, the 
communication cost is negligible compared to other computations performed on each process.

### 2024-05-21 paper/examples.tex
New tweaks based on Alex's comments

\subsubsection{Online statistics}
\label{sec:online_stats}
Having specified the use of the \texttt{online} recorder in our call to \texttt{pigeons()}, 
we can output some basic summary statistics of the marginals of our target distribution. 
For instance, it is straightforward to estimate the mean and variance of each 
of the marginals of the target with 
\begin{lstlisting}[language=Julia]
using Statistics
mean(pt); var(pt)
\end{lstlisting}
Other constant-memory statistic accumulators are made available in the 
OnlineStats.jl \cite{day2020onlinestats} package. 
To add additional constant-memory statistic accumulators, we 
can register them via \texttt{Pigeons.register\_online\_type()}, as described in our 
online documentation. For instance, we can also compute constant-memory estimates 
of extrema of our distribution.

>>>>

\subsubsection{Online statistics}
\label{sec:online_stats}
Pigeons has facilities to support cases requiring large memory. For instance, we
allow for the computation of online statistics, as well as off-memory sample storage.
Having specified the use of the \texttt{online} recorder in our call to \texttt{pigeons()}, 
we can output some basic summary statistics of the marginals of our target distribution. 
For instance, it is straightforward to estimate the mean and variance of each 
of the marginals of the target with 
\begin{lstlisting}[language=Julia]
using Statistics
mean(pt); var(pt)
\end{lstlisting}
Other constant-memory statistic accumulators are made available in the 
OnlineStats.jl \cite{day2020onlinestats} package. 
To add additional constant-memory statistic accumulators, we 
can register them via \texttt{Pigeons.register\_online\_type()}, as described in our 
online documentation. For instance, we can also compute constant-memory estimates 
of extrema of our distribution.
